Study of the hadronization of charged pions

### 2016-09-27 AnalysisNote.tex
Edit on the part 1

\begin{abstract}
Hadronization is the process by which energetic quarks transform into 
colorless hadrons. The process is non-perturbative, therefore, only 
qualitative understanding can be achieved, and using phenomenological models 
is necessary. It is believed that hadronization has two distinct phases: 
the quark propagation phase, where the quark emit gluons until it neutralizes 
its color by forming a pre-hadron and the hadron formation phase, where the 
pre-hadron evolves into a hadron. Those process are respectively characterized 
by the production time and the formation time. By comparing hadron production 
on different nuclei, one can measure variables with different sensitivities 
to the hadronization phases, the two principal variables are: the transverse 
momentum broadening, that is believed to be tightly connected to quark energy 
loss and the multiplicity ratio, which is a measure of the hadron attenuation 
in the medium.

The goals of the experiment E-02-104 are to study both quark propagation
and hadronization dynamic using various nuclear target.
In this analysis we focus on two particles, the negative pion and the proton.
Our goals are to constrain the theoritical models and extract the production
time using precise multidimensional analysis of $\pi^+$ and $\pi^-$ to enlight possible
new physical processes with a precise analysis of proton production in nuclei.

@@

\bibliography{hadro}
\bibliographystyle{plain}

>>>>

\begin{abstract}
Hadronization is the process by which energetic quarks transform into 
colorless hadrons. The process is non-perturbative, therefore, only 
qualitative understanding can be achieved, and using phenomenological models 
is necessary. By comparing hadron production 
on different nuclei, one can measure variables with different sensitivities 
to the hadronization phases, the two principal variables are: the transverse 
momentum broadening, that is believed to be tightly connected to quark energy 
loss and the multiplicity ratio, which is a measure of the hadron attenuation 
in the medium.

The goals of the experiment E-02-104 are to study both quark propagation
and hadronization dynamic using various nuclear target.
In this analysis we focus on two particles, the positive and negative pions
to constrain theoritical models using precise multidimensional analysis.

@@

\bibliography{hadro}
\bibliographystyle{ieeetr}

### 2016-09-27 intro.tex
Edit on the part 1

The fragmentation process by which parton fragment into hadrons is 
experimentally well studied and fragmentation functions fitted from 
data are measured on a large kinematic and for numerous particles
\cite{AKK}.
Nevertheless, the hadron formation is a non perturbative process and can not 
be theoritically described, only a global picture 
(see figure \ref{fig:hadro}) can be drawn for
energies above the resonance region. In deep inelastic scattering the
virtual photon interact with a quark which propagate quasi-free emitting
gluons during the 
so-called production time. After neutralizing its color the quark become 
a pre-hadron which will eventualy form a hadron after the formation time.
On nuclear target the process of hadronization occurs, at least partly, in 
nuclear matter, then one can extract the production time using the different 
behaviour of quarks and colorless pre-hadrons in medium. The measurement of
pions is especially suited to lead that kind of analysis because they are easy
to detect and abundant in deep inelastic scattering.

@@

In order to constrain the existing models we extract our results
using multi-dimensional binning. Another way to constrain the models will be
the observation of baryon hadronization (here protons), which behaviour is
not understood yet. 
The importance of the EG2 data then come from its high statistics which permit, 
this multi-dimensional measurement of both pions and protons on a 
large kinematic range.

@@

In this analysis we use the following semi-inclusive DIS variables:
\begin{itemize}
 \item $\nu = E_i - E_f$ with $E_i$ the beam energy and $E_f$ the scattered electron energy,
 \item $Q^2 = 4 E_i E_f \sin ^2(\theta_e / 2)$ with $\theta_e$ the scattered electron angle with the beam line,
 \item $x_{Bj} = {{Q^2} \over {2 M_n \nu}}$ with $M_n$ the nucleon mass,
 \item $W^2 = M_n^2 - Q^2 + 2 M_n \nu$
 \item $z_h = E_h / \nu$
 \item $P_T^2$ which is transverse momentum according to the virtual photon
 \item $\phi_h$ the angle between the electronic plan (contain initial and scattered electrons) and the hadronic plan (contain virtual photon and hadron)
 \item add xf
\end{itemize}

We define the multiplicity ratio as:
\begin{equation}
R_A^h (Q^2,\nu,z_h,P_T^2) = {{N_A^h (Q^2,\nu,z_h,P_T^2) / N_A^e (Q^2,\nu)} 
                       \over {N_D^h (Q^2,\nu,z_h,P_T^2) / N_D^e (Q^2,\nu)}}
\end{equation}
with $N^e_t$ the number of electron measured from a target t and $N_t^h$ the number
of semi-inclusive hadrons h (detection of both an electron and a hadron) from the
target~t. The multiplicity ratio represents the attenuation of the hadron h in a 
nuclear target~A.
\newline

We define the transverse momentum broadening as:
\begin{equation}
\Delta \langle P_T^2 \rangle = \langle P_T^2 \rangle_A - \langle P_T^2 \rangle_D
\end{equation}
with $\langle P_T^2 \rangle_t$ the mean transverse momentum measured in a target t.

Finally a less used observable to carachterise the momentum broadening is the ratio
of mean transverse momentum:

\begin{equation}
\langle P_T^2 \rangle_A / \langle P_T^2 \rangle_D
\end{equation}

@@

In \cite{Wang:2002ri}, E.~Wang and X.-N.~Wang, describe HERMES data using only energy 
loss from the parton, the suppression observed is then due to the fact that a 
lower energy quark will fragment into a lower number of hadrons and a lower 
z. That kind of model is used in both RHIC and nDIS and permit a commom 
interpretation of hadron suppression in nuclear matter. Lot of different
calculations of the energy loss exist, the main parameter for those is $\hat q$
(GeV$^2$ fm$^{-1}$) which is the transverse momentum per unit of length of a quark after propagating
in nuclear matter, this value is related to the $P_T$ 
broadening observables. The two main difficulties for the quark energy loss
models are to have coherent picture of both multiplicity ratios and $P_T$ 
broadening because usually the $\hat q$ needed to reproduce multiplicity ratio
is bigger than the one necessary to describe transverse momentum broadening.

The GiBUU model \cite{Gallmeister:2007an} is a transport model based on Boltzmann equation
using hadronic and pre-hadronic interaction in the nuclear matter, with no
quark energy loss involved. This model reproduce very well most of the hadrons, 
surprisingly, it does not reproduce protons
data from HERMES correctly, it is surprising in the sense that the proton
case should be well handle by this kind of models. Another problem 
is that the $\Delta \langle P_T^2 \rangle$ variable is not yet described by 
hadron absorption models.

@@

To conclude it is important to point out that no consensus is reached on which 
mecanisms are dominant and it is important to perform precise measurment with
the right observables to solve this problem.

@@

Hadron multiplicity ratios in nuclei were measured in numerous lepton 
facilities: L.S.~Osborne {\it et al.} \cite{Osborne:1978ai} at SLAC,
L.~Hand {\it et al.} \cite{Hand:1978tx} and the E665 collaboration \cite{Adams:1994ri} at FNAL
and the European Muon Collaboration \cite{Arvidson:1984fz,Ashman:1991cx} at CERN. Those 
measurements revealed the general picture; we observe a sizeable suppression when
looking at hadron production in nuclei, this suppression is lower at higher
$\nu$ and lower z, it can even inverse and increase at very low z.

The figures \ref{fig:her1}, \ref{fig:her2} and \ref{fig:her3} shows a sample of 
more recent data from HERMES collaboration \cite{Airapetian:2007vu}, numerous 
hadrons are studied individually and new variables linked with transverse 
momentum are used on top of the usual multiplicity ratio.

\begin{figure}[htbp]
\centering
\includegraphics[width=14cm] {fig/Hermes/pi0hermes.png} 
\caption {Multiplicity ratio of $\pi^0$ in function of various kinetic 
variables from HERMES collaboration \cite{Airapetian:2003mi}}
\label{fig:her1}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=14cm] {fig/Hermes/hermes1.png} 
\caption {Multiplicity ratio of $\pi^+$, K$^+$, protons and anti-protons in 
function of various kinetic variables from HERMES collaboration \cite{Airapetian:2007vu}}
\label{fig:her2}
\end{figure}

>>>>

The hadronization process by which parton fragment into hadrons is 
experimentally well studied and fragmentation functions fitted from 
data are measured on a large kinematic range~\cite{Albino:2008fy}.
Nevertheless, the hadron formation is a non perturbative process and can not 
be theoritically described, only a global picture 
(see figure \ref{fig:hadro}) can be drawn for
energies above the resonance region. In deep inelastic scattering the
virtual photon interact with a quark which propagate quasi-free emitting
gluons during the 
so-called production time. After neutralizing its color the quark become 
a pre-hadron which will eventualy form a hadron after the formation time.
On nuclear target the process of hadronization occurs, at least partly, in 
nuclear matter, then one can extract the production time using the different 
behaviour of quarks and colorless pre-hadrons in medium. The measurement of
pions is especially suited to lead that kind of analysis because they are easy
to detect and abundant in deep inelastic scattering.

@@

In order to constrain the existing models we extract our results
using tight multi-dimensional binning. 
The importance of the EG2 data comes from its high statistics which permit, 
this multi-dimensional measurement of both pions on a 
large kinematic range.

@@

In this analysis we use the following semi-inclusive DIS variables:
\begin{itemize}
 \item $\nu = E_i - E_f$ with $E_i$ the beam energy and $E_f$ the scattered electron energy,
 \item $Q^2 = 4 E_i E_f \sin ^2(\theta_e / 2)$ with $\theta_e$ the scattered electron angle with the beam line,
 \item $x_{Bj} = {{Q^2} \over {2 M_n \nu}}$ with $M_n$ the nucleon mass,
 \item $W^2 = M_n^2 - Q^2 + 2 M_n \nu$
 \item $z_h = E_h / \nu$
 \item $P_T^2$ which is transverse momentum according to the virtual photon
 \item $\phi_h$ the angle between the electronic plan (contain initial and scattered electrons) and the hadronic plan (contain virtual photon and hadron)
 \item $x_F = P_L/P_L^{max}$ the fraction of longitudinal momentum carried by the hadron calculated in respect to the virtual photon direction in the laboratory frame.
\end{itemize}
%TODO Add t

We define the multiplicity ratio as:
\begin{equation}
R_A^h (Q^2,\nu,z_h,P_T^2) = {{N_A^h (Q^2,\nu,z_h,P_T^2) / N_A^e (Q^2,\nu)} 
                       \over {N_D^h (Q^2,\nu,z_h,P_T^2) / N_D^e (Q^2,\nu)}}
\end{equation}
with $N^e_A$ the number of electron measured from a target $A$ and $N_A^h$ the number
of semi-inclusive hadrons $h$ (detection of both an electron and a hadron) from the
target~$A$. The multiplicity ratio represents the attenuation of the hadron $h$ in a 
nuclear target~$A$.
\newline

We define the transverse momentum broadening as:
\begin{equation}
\Delta \langle P_T^2 \rangle = \langle P_T^2 \rangle_A - \langle P_T^2 \rangle_D
\end{equation}
with $\langle P_T^2 \rangle_A$ the mean transverse momentum measured in a target $A$.

@@

In \cite{Wang:2002ri}, E.~Wang and X.-N.~Wang, describe HERMES data using only energy 
loss from the parton, the suppression observed is then due to the fact that a 
lower energy quark will fragment into a lower number of hadrons and a lower 
$z$. That kind of model is used in both RHIC and nDIS and permit a commom 
interpretation of hadron suppression in nuclear matter. Lot of different
calculations of the energy loss exist, the main parameter for those is $\hat q$
(GeV$^2$ fm$^{-1}$) which is the transverse momentum per unit of length of a 
quark after propagating
in nuclear matter, this value is related to the $P_T$ 
broadening observables. The main difficulty for the quark energy loss
models is to have coherent description of both multiplicity ratios and $P_T$ 
broadening. Recent models find a $\hat q$ to reproduce multiplicity ratio
which is bigger than the one necessary to describe transverse momentum broadening.
% TODO Add a ref

The GiBUU model \cite{Gallmeister:2007an} is a transport model based on Boltzmann equation
using hadronic and pre-hadronic interaction in the nuclear matter, with no
quark energy loss involved. This model reproduce very well most of the hadrons, 
however the $\Delta \langle P_T^2 \rangle$ variable is not described at all by 
this hadron absorption models.

@@

To conclude it is important to point out that no consensus is reached on which 
mecanisms are dominant. It is therefore important to perform precise measurements with
the right observables to solve this problem.

@@

Hadron multiplicity ratios in nuclei were measured in numerous lepton 
facilities: L.S.~Osborne {\it et al.} \cite{Osborne:1978ai} at SLAC,
L.~Hand {\it et al.} \cite{Hand:1978tx} and the E665 collaboration \cite{Adams:1994ri} at FNAL
and the European Muon Collaboration \cite{Arvidson:1984fz,Ashman:1991cx} at CERN. Those 
measurements revealed the general features of hadronization
in nuclei: we observe a sizeable suppression when
looking at hadron production in nuclei, this suppression is lower at higher
$\nu$ and lower $z$, it can even inverse and be an increased yield at very low $z$.

The figures \ref{fig:her1}, \ref{fig:her2} and \ref{fig:her3} shows a sample of 
the most recent data from HERMES collaboration \cite{Airapetian:2007vu}, numerous 
hadrons are studied individually and new variables linked with transverse 
momentum are used on top of the usual multiplicity ratio.

\begin{figure}[htbp]
\centering
\includegraphics[width=14cm] {fig/Hermes/pi0hermes.png} 
\caption {Multiplicity ratio of $\pi^0$ as a function of various kinetic 
variables from HERMES collaboration \cite{Airapetian:2003mi}}
\label{fig:her1}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=14cm] {fig/Hermes/hermes1.png} 
\caption {Multiplicity ratio of $\pi^+$, K$^+$, protons and anti-protons as a 
function of various kinetic variables from HERMES collaboration \cite{Airapetian:2007vu}}
\label{fig:her2}
\end{figure}
